---
title: "Homework 1"
author: "Rudy Martinez"
date: "10/04/2020"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

# Set your theme here
theme_set(theme_minimal())

# Create a duplicate copy of diamonds data here
diam_1 = diamonds

# Create Color Palettes
colors_1 = c("Fair" = "#99B898",
              "Good" = "#FECEAB", 
              "Very Good" = "#FF847C", 
              "Premium" = "#E84A5F", 
              "Ideal" = "#2A363B")

colors_2 = c("I1" = "purple4",
             "SI2" = "mediumpurple4",
             "SI1" = "steelblue4",
             "VS2" = "darkslategray4",
             "VS1" = "lightseagreen",
             "VVS2" = "mediumseagreen",
             "VVS1" = "olivedrab3",
             "IF" = "gold")
```



In this homework, you will use `diamonds` data set from `ggplot2` package. It is automatically loaded when you execute `library(ggplot2)` so you don't have to separately load it. Please create a duplicate of `diamonds` and use that for homework. This will avoid corrupting the original data set.

Make sure that you understand the variables in the data by checking the help file by executing `help(diamonds)` in the console. The data has five variables that measure the dimensions. 

## Instructions

1. You are expected to recreate each plot **exactly** as shown in this homework.
2. The objective of this homework is to help you develop fine grain control over `ggplot2`. As such, please refrain from exercising artistic freedom!
3. All the plots use `theme_minimal()`. You can set this as your default theme by adding this line in the `setup` chunk **after** you load `ggplot2` library:
`theme_set(theme_minimal())`


## Q1 (3 points)

Recreate the following graph. The parameter that controls transparency is set at 0.3. You need not get exactly the same colors but they must be **discrete** and should not follow a color gradient.


```{r Q1}
ggplot(diam_1, aes(carat,price)) +
  geom_point(aes(fill = as.character(clarity)),
             color = "black",
             alpha = 0.3,
             shape = 21,
             size = 1.5) +
  labs(title = "Scatterplot of Diamond Prices", x = "Diamond Carats", y = "Diamond Price") + 
  scale_y_continuous(labels = scales::dollar_format(prefix = "$"))
```


## Q2 (2 points)

The previous graph looks cluttered. So you decided to use facets instead. Recreate the following graph:


```{r Q2}
ggplot(diam_1, aes(carat,price)) +
  geom_point(aes(fill = as.character(clarity)),
             color = "black",
             alpha = 0.3,
             shape = 21,
             size = 1.5,
             show.legend = FALSE) +
  labs(title = "Scatterplot of Diamond Prices", x = "Diamond Carats", y = "Diamond Price") + 
  scale_y_continuous(labels = scales::dollar_format(prefix = "$")) +
  facet_wrap(~clarity, nrow = 2, ncol = 4)
```

## Q3 (5 points)

Next, you want to know whether the price of diamonds depends on table and depth. Note the line types. Recreate the following graphs:


```{r Q3.1, message=FALSE}
ggplot(diam_1, aes(table, price)) + 
  geom_smooth(method = "lm",
              formula = y ~ x,
              alpha = 0.3,
              color = "red",
              linetype = "dashed",
              fill = "grey47",
              show.legend = FALSE) + 
  scale_x_continuous(breaks = seq(0,100, by = 25),
                     limits = c(0,100)) + 
  scale_y_continuous(breaks = seq(0,10000, by = 5000)) +
  labs(x = "Table", y = "Price")
```


```{r Q3.2, message=FALSE}
ggplot(diam_1, aes(depth, price)) + 
  geom_smooth(method = "lm",
              formula = y ~ x,
              alpha = 0.3,
              color = "white",
              linetype = "dotdash",
              fill = "grey47",
              show.legend = FALSE) + 
  scale_x_continuous(breaks = seq(0,80, by = 10),
                     limits = c(0,80)) + 
  scale_y_continuous(breaks = seq(3000,5000, by = 500)) +
  labs(x = "Depth", y = "Price")
```



## Q4 (5 points)

Recreate each of the following graphs for data exploration:


```{r Q5.1, message=FALSE}
ggplot(diam_1, aes(x*y*z, price)) +
  geom_point(aes(color = cut)) +
  scale_fill_manual(values = colors_1) + 
  scale_color_manual(values = colors_1) +
  labs(x = "x * y * z", y = "price")
```


```{r Q5.2}
ggplot(diam_1, aes(price)) +
  geom_histogram(bins = 75, fill = "grey35", color = "white") +
  scale_x_continuous(labels = scales::dollar_format(prefix = "$"),
                     breaks = seq(0,15000, by = 5000)) +
  scale_y_continuous(labels = scales::comma_format(),
                     breaks = seq(0,7000, by = 2000))
```



```{r Q5.3}
ggplot(diam_1, aes(clarity)) +
  geom_bar(fill = colors_2, color = "red") +
  scale_y_continuous(breaks = seq(0,10000, by = 5000))
```


```{r Q5.4}
ggplot(diam_1, aes(cut, depth)) + 
  geom_violin(color = "blue") +
  geom_jitter(color = "red",  alpha = 0.03)
```


```{r Q5.5, message=FALSE}
ggplot(diam_1, aes(x, price)) +
  geom_smooth(method = "lm",
              color = "green",
              show.legend = FALSE) + 
  geom_smooth(fill = "white",
              show.legend = FALSE)
```


