---
title: "Superimposing Summary by Group"
author: "Ashwin Malshé"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

Often we need to plot summary statistics for each group on top of the raw data. For this, we need two data sets.

In the following example, I will use mpg data set and plot average of `hwy` on top of a scatterplot. Here the average is NOT by any group and instead it's overall average.

```{r}
hwymean1 <- summarize(mpg, avghwy = mean(hwy))
```

```{r}
print(hwymean1)
```

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_point() +
  geom_hline(aes(yintercept = avghwy), 
             data = hwymean1, 
             color = "red") + 
  theme_minimal()
```


What if we want to plot average values of each group?

We need to have a data frame of average values of each group. It can be done in a similar fashion as the previous graph. However, the code will be cleaner if we create a data frame separately.

```{r}
hwymean2 <- mpg %>%
            group_by(class) %>%
              summarize(avghwy = mean(hwy),
                        .groups = "drop") # This will ungroup the data
```


Now we can use the same code as before

```{r}

ggplot(mpg, aes(class, hwy)) +
  geom_point() +
  geom_point(aes(x = class, y = avghwy), 
             data = hwymean2, 
             color = "red", 
             size = 4,
             alpha = 0.5) + 
  theme_minimal()
```

Finally, plot the overall mean on the graph as well

```{r}

ggplot(mpg, aes(class, hwy)) +
  geom_point() +
  geom_point(aes(y = avghwy), 
             hwymean2, 
             color = "red", 
             size = 4,
             alpha = 0.5) +
  geom_text(aes(label = round(avghwy, 2), y = avghwy), 
            data = hwymean2, 
            color = "red",
            hjust = 1.3
            ) +
  geom_hline(aes(yintercept = avghwy), 
             data = hwymean1, 
             color = "blue") +
  geom_text(aes(x = "compact", y = 22, label = round(avghwy,2)),
            data = hwymean1, 
            color = "blue", 
            hjust = 1.5) + 
  theme_minimal()

# "compact" is used to place the value for the overall average above "compact" group. It is placed here because it is more visible. It can be anywhere. 
# y = 22 is used to placed the value for the overall average in line with y = 22 to make the value more visible. It can be anywhere.
# hjust argument (horizontal justification) enables you to adjust the left to right position of the label.
```

If you want to make it easier for people to read, perhaps ordering the vehicle class by average hwy will help

```{r}
m1 = mpg %>% 
  inner_join(hwymean2, by = "class") %>% 
  mutate(class = reorder(class, avghwy))

```

```{r}

ggplot(m1, aes(class, hwy)) +
  geom_point() +
  geom_point(aes(y = avghwy), 
             data = hwymean2, 
             color = "red", 
             size = 4) +
  geom_text(aes(label = round(avghwy, 2), y = avghwy), 
            data = hwymean2, 
            color = "red", 
            hjust = 1.3) +
  geom_hline(aes(yintercept = avghwy), 
             data = hwymean1, 
             color = "blue") +
  geom_text(aes(x = "compact", y = 22, label = round(avghwy,2)),
            data = hwymean1, 
            color = "blue", 
            hjust = 1.5) 

```
