---
title: "ggplot2-basics"
author: "Ashwin Malshe"
date: "22 September 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## The structure of a `ggplot2` plot

```{r eval = FALSE}

ggplot(data = ,
       mapping = aes()) +
  geom_*() +
  scale_*() +
  labs() + 
  facet_*() +
  theme()

```


## `mpg` dataset

Let's start with a dataset named `mpg` that comes bundled with `ggplot2`

Get the variable names

```{r}
names(mpg)
```

Check the `class` of each variable in `mpg`

```{r}
sapply(mpg, class)
```


## `ggplot()` function


`ggplot()` is the key function in `ggplot2`. If you execute the following chunk, you will get a gray box as an output. This is a standard output from `ggplot()` in the absence of any information on the data.

```{r}
ggplot()
```


`ggplot()` uses `data` and `mapping` arguments to create the base layer for a plot. Giving `data` alone won't output anything informative.


```{r}
ggplot(data = mpg)
```

We need to provide the mapping from the data to aesthetics. This is not always required because some `geoms` will have their own default statistics mapping to the aesthetic. Consider this as a case where a `geom` creates a variable from the data and then uses that to make the plot.

In the next chunk, we provide the mapping from `cty` to the X axis and `hwy` to the Y axis. This will lead to a more informative plot compared to a large gray box!


```{r}
ggplot(data = mpg,
       mapping = aes(x = cty, y = hwy))
```


We still don't know what we want to show on the plot. However, now we can start to think about the possible visualizations. Both `cty` and `hwy` are continuous variables. So making a scatterplot to show their relationship will be informative.

We do that by **adding** `geom_point()` to `ggplot()`

Note two important aspects here:

1. We add a layer to a `ggplot` by using a `+` sign. This is not a ` %>% `! People make this mistake very often so be careful here.

2. We don't provide any `data` or `mapping` to `geom_point()`. This is because all the layers inherit these arguments from `ggplot()`. So you don't have to keep on specifying them in each layer.

*Important: You can specify `data ` and `mapping` inside any `geom_`. It will not be passed on to any other layer. It will remain confined to that layer alone.*


```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point()
```

This is the same as above:

```{r}
ggplot() +
  geom_point(data = mpg, mapping = aes(x = cty, y = hwy))
```


Note that we can add another layer and it will inherit the data and mapping from `ggplot` again.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() +
  geom_line()
```

But if `ggplot()` doesn't specify them, subsequent layers don't inherit anything from another layers. So in the code below, the line doesn't show.

```{r}
ggplot() +
  geom_point(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_line()
```

Adding the title, subtitle, and the caption is really easy.


```{r}
ggplot(mpg, aes(x = cty, y = hwy)) +
  geom_point() +
  labs(title = "City Miles vs Highway Miles",
       subtitle = "Based on the data from 1999 and 2008",
       caption = "Created by Ashwin Malshe using mpg dataset.
       mpg dataset is included in ggplot2 package.",
        x = "miles per gallon in city driving",
        y = "miles per gallon in highway driving")
```


We can show subgroups by using various aesthetics. For example, we can color the points based on the drive of the cars.


```{r}
ggplot(mpg, aes(x = cty, y = hwy, color = drv)) +
  geom_point() +
  labs(title = "City Miles vs Highway Miles",
       subtitle = "Based on the data from 1999 and 2008",
       caption = "Created by Ashwin Malshe using mpg dataset.")
```


Finally, instead of showing the groups separately may make sense in some cases. This code will split the data into 3 groups based on the value of `drv` and then create a plot for each subgroup. This is known as a facet plot.


```{r}
ggplot(mpg, aes(x = cty, y = hwy, color = drv)) +
  geom_point() +
  facet_wrap(~ drv)
```

