---
title: 'Session 2'
author: "Rudy Martinez"
date: "9/1/2020"
output: html_document
---

This next R code chunk will output a vector.

```{r Vector 1}
vec1 = c("a", "b", "c", "d", "e", "f")
print(vec1)
```

This next R code chunk will output a numeric vector.

```{r Vector 2}
vec2 = c(3.14159, 1.414, 5.678, 9.01)
print(vec2)
```

## Matrix
A matrix is a two-dimensional vector.

```{r Matrix 1}
m1 = matrix(c(3.14159, 1.414, 5.678, 9.01),
            nrow = 2,
            ncol = 2)
print(m1)
```

```{r Matrix 2}
m2 = matrix(vec2, nrow = 2)
print(m2)
```

## Data Frame
```{r Data Frame}
d1 = data.frame(name = c("Rudy", "John", "Andrew"), 
                age = c(20, 21, 22),
                male = c(TRUE, TRUE, TRUE),
                female = c(FALSE, FALSE, FALSE))
print(d1)
```

## Subsetting Data Frame
Print first column of d1

```{r Subsetting Data Frame 1}
print(d1[, 1])

```

Output the intersection of second row and third column

```{r Subsetting Data Frame 2}
print(d1[2,3])
```

Print the column names of the data frame
```{r Column Names}
names(d1)
```

Print the first two rows and first two columns of d1
```{r First Two Rows / Columns 1}
print(d1[c(1,2),c(1,2)])
```

OR

```{r First Two Rows / Columns 2}
print(d1[1:2, 1:2])
```

Print 1st and 3rd row and 1st and 3rd columns
```{r First and Third Row / Column}
print(d1[c(1,3), c(1,3)])
```

## List
```{r List}
list1 = list(v1 = vec1, 
             v2 = vec2,
             matrix = m1,
             dataframe = d1)
```

