---
title: "Session Advanced Topics"
author: "Rudy Martinez"
date: "9/8/2020"
output: html_document
---

## Libraries
```{r Libraries}
library(ggplot2)
library(tidyverse)
library(dplyr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo= TRUE)
```

### Exercise 1
Use the data frame `d1` and keep all the rows where age `age` is <= 50

For this task, we need to use `filter()` from `diplyr`
```{r Exercise 1}
d2 = filter(d1, age <= 50)
d2
```

### Exercise 2
Use the data frame `d1` and keep all the rows where age `age` is <= 21 and `male` is TRUE
```{r Exercise 2}
d3 = filter(d1, age <= 21 | male == TRUE)
d3
```

### Exercise 3
From `d1`, keep only the first two columns labeled `name` and `age`
```{r Execerise}
select(d1, name, age)
```

### Exercise 4
Extract all the columns with names starting with letter a
```{r Exercise 4}
select(d1, starts_with(("a")))
```

### Exercise 5
Extract all the numeric columns
```{r Exercise 5}
select(d1, is.numeric)
```
### Exercise 6
Rename `name` as `first_name` and rename `age` as `age_in_years`
```{r Exercise}
rename(d1, 
       first_name = name,
       age_in_years = age)
```

