---
title: "gganimate"
author: "Ashwin Malshe"
date: "06 October 2020"
output: html_document
---

Install `gganimate` package by executing the following command in in the console below.

`install.packages("gganimate")`

```{r setup, include=FALSE}
library(ggplot2)
library(gganimate)
library(dplyr)
```

```{r}
plot1 <- ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  theme_minimal()
```

In case you want to take a look:
```{r}
print(plot1)
```


```{r}
anim1 <- plot1 +
  transition_states(class,
                    transition_length = 2,
                    state_length = 1)
```

```{r}
print(anim1)
```


```{r}
anim1 + labs(title = 'Class: {closest_state}')
```


## Add color to the plot

```{r}
plot2 <- ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point(show.legend = FALSE) +
  theme_minimal()
```


```{r}
print(plot2)
```


```{r}
anim2 <- plot2 +
  transition_states(class,
                    transition_length = 2,
                    state_length = 1) +
  labs(title = 'Class: {closest_state}')
```

```{r}
print(anim2)
```


Remove the jerkiness by using `group`.

```{r}
plot3 <- ggplot(mpg, 
                aes(displ, hwy, 
                    color = class, group = 1)) +
  geom_point(show.legend = FALSE) +
  theme_minimal()
```

```{r}
anim3 <- plot3 +
  transition_states(class,
                    transition_length = 2,
                    state_length = 1) +
  labs(title = 'Class equals {closest_state}')
```

```{r}
animate(anim3, nframes = 50)
```

```{r}
anim_save(here::here("Images", "anim_20201002.gif"))
```


## Resources

Unfortunately the package doesn't have a lot of examples. The developer of the package Thomas Pedersen also manages `ggplot2` so he is quite busy. However, here are two good resource - 

The official package website for `gganimate`:

https://gganimate.com

Go to the bottom of this page for sveral examples:

https://github.com/ropenscilabs/learngganimate
