---
title: "Top 10 Billionaires"
author: "Ashwin Malshe"
date: "06 October 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggthemes)
# pacman::p_load(ggplot2, dplyr, ggthemes)
```

Get the data

```{r}
top10 <- readRDS(gzcon(url("https://github.com/ashgreat/datasets/blob/master/top10_billionaires.RDS?raw=true"))) %>% 
  filter(year != 2018)
```

Take a peek

```{r}
head(top10, 10)
```


### The following code will give error. Why?


```{r eval=FALSE}
ggplot(top10, aes(x = name, y = netWorth, fill = year)) +
  geom_bar() +
  labs(x = "Name", 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth")
```


Adding identity (or you can use `geom_col`)

```{r}
ggplot(top10, aes(x = name, y = netWorth, fill = year)) +
  geom_bar(stat = "identity") +
  labs(x = "Name", 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth")

```

The position of the bars is stacked. We want them to be dodged

Play around with the widths

```{r}
ggplot(top10, aes(x = name, y = netWorth, fill = year)) +
  geom_bar(stat = "identity", 
           position = position_dodge(0.9)) +
  labs(x = "Name", 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth")

```


Why is the dodging not working?

In order to godge the bars, the `fill` argument should be mapped to a categorical variable. So I will create a new categorical variable.

```{r}
top10$year2 <- factor(top10$year)
```



```{r}
ggplot(top10, aes(x = name, y = netWorth, fill = year2)) +
  geom_bar(stat = "identity", 
           position = position_dodge(0.9)) +
  labs(x = "Name", 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth")

```



Change the legend title to "Year"

```{r}
ggplot(top10, aes(x = name, 
                  y = netWorth, 
                  fill = year2)) +
  geom_bar(stat = "identity", 
           position = position_dodge(width = 0.9)) +
  labs(x = "Name", 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth",
       fill = "Year")

```


The names appear in alphabetical order. Either we can recode the names to be factors or we can change the ordering based on the ranking in ggplot.


```{r}
ggplot(top10, aes(x = reorder(name, rank),
                  y = netWorth, 
                  fill = year2)) +
  geom_bar(stat = "identity",
           position = position_dodge(width = 0.9)) +
  labs(x = "Name",
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth",
       fill = "Year")
```

Something is still missing here. Any idea what's going on?


```{r}

top2017 <- top10 %>% 
  filter(year == 2017) %>% 
  select(name, rank) %>% 
  rename(rank2017 = rank)  

top10 <- inner_join(top10, top2017, by = "name")

# print column names
colnames(top10)

top10$name <- reorder(top10$name, top10$rank2017)

```

Also, split the names on two lines so that we can render them nicely

```{r}
top10$name2 = gsub(" ", " \n", top10$name)
top10$name2 = reorder(top10$name2, top10$rank2017)

```


```{r}
ggplot(top10, aes(x = name2, 
                  y = netWorth, 
                  fill = year2)) +
  geom_bar(stat = "identity",
           position = position_dodge(width = 0.9)) +
  labs(x = "Name",
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth",
       fill = "Year")
```


## Adjust the X axis ticks by changing the arguments in `element_text()`

```{r}
ggplot(top10, aes(x = name2, 
                  y = netWorth, 
                  fill = year2)) +
  geom_bar(stat = "identity",
           position = position_dodge(width = 0.9)) +
  labs(x = "Name", 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth", 
       fill = NULL) +
  theme(axis.text.x = element_text(angle = , hjust =, size = ))
 
```

  
## Add commas to Y axis labels

```{r}
ggplot(top10, aes(x = name2, 
                  y = netWorth, 
                  fill = year2)) +
  geom_bar(stat = "identity",
           position = position_dodge(width = 0.9)) +
  labs(x = "Name",
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth",
       fill = NULL) +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 0.6)) +
  scale_y_continuous(labels = scales::comma)

```


### Label the bars

```{r}
ggplot(top10, aes(x = name,
                  y = netWorth, 
                  fill = year2)) +
  geom_bar(stat = "identity",
           position = position_dodge(width = 0.9)) +
  labs(x = "Name", 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth", 
       fill = NULL) +
  scale_y_continuous(labels = scales::comma) +
  geom_text(aes(label = netWorth)) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 0.6))

```


The labels are clearly badly rendered
We need to dodge them the same way we dodged the bars

```{r}
ggplot(top10, aes(x = name, 
                  y = netWorth, 
                  fill = year2)) +
  geom_bar(stat = "identity", 
           position = position_dodge(width = 0.9)) +
  labs(x = "Name", 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth", 
       fill = NULL) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 0.6)) +
  scale_y_continuous(labels = scales::comma) +
  geom_text(aes(label = netWorth),
            position = position_dodge(width = 0.9))
```

Looks better but still not the best
Rotate the labels by 90 degrees

```{r}
ggplot(top10, aes(x = name, y = netWorth, fill = year2)) +
  geom_bar(stat = "identity", 
           position = position_dodge(width = 0.9)) +
  labs(x = "Name", 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth", 
       fill = NULL) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 0.6)) +
  scale_y_continuous(labels = scales::comma) +
  geom_text(aes(label = netWorth),
            position = position_dodge(width = 0.9),
            angle = 90)

```


### Push the labels inside the bars


```{r}
ggplot(top10, aes(x = name, 
                  y = netWorth, 
                  fill = year2)) +
  geom_bar(stat = "identity", 
           position = position_dodge(width = 0.9)) +
  labs(x = "Name", 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth", 
       fill = NULL) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 0.6)) +
  scale_y_continuous(labels = scales::comma) +
  geom_text(aes(label = netWorth),
            position = position_dodge(width = 0.9),
            angle = 90,
            hjust = 1.1)

```


### Change the color of the labels

```{r}
ggplot(top10, aes(x = name, 
                  y = netWorth, 
                  fill = year2)) +
  geom_bar(stat = "identity", 
           position = position_dodge(width = 0.9)) +
  labs(x = "Name", 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth", 
       fill = NULL) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 0.6)) +
  scale_y_continuous(labels = scales::comma) +
  geom_text(aes(label = netWorth),
            position = position_dodge(width = 0.9),
            angle = 90,
            hjust = 1.1,
            color = "white")

```

## Make this look cool using premade themes

Go to the top of the file and insert `library(ggthemes)` in the setup chunk.

```{r}
ggplot(top10, aes(x = name2, 
                  y = netWorth, 
                  fill = year2)) +
  geom_bar(stat = "identity", 
           position = position_dodge(width = 0.9)) +
  labs(x = NULL, 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth", 
       fill = NULL) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 0.6)) +
  scale_y_continuous(labels = scales::comma) +
  geom_text(aes(label = netWorth),
            position = position_dodge(width = 0.9),
            angle = 90,
            hjust = 1.1,
            color = "white") +
  theme_economist()
```



Usually themes have their own parameters and they can override our themes.
So let's put our theme the last.

```{r}
ggplot(top10, aes(x = name2, 
                  y = netWorth, 
                  fill = year2)) +
  geom_bar(stat = "identity", 
           position = position_dodge(width = 0.9)) +
  labs(x = NULL, 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth", 
       fill = NULL) +
  scale_y_continuous(labels = scales::comma) +
  geom_text(aes(label = netWorth),
            position = position_dodge(width = 0.9),
            angle = 90,
            hjust = 1.1,
            color = "white") +
  theme_economist() +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 0.5, 
                                   vjust = 0.7))

```

Save the file to your hard disk

```{r eval=FALSE}
ggsave("Top_10_Billionaires.png", 
       width = 9, 
       height = 6)
```

### Try a different theme

```{r}
ggplot(top10, aes(x = name2, 
                  y = netWorth, 
                  fill = year2)) +
  geom_bar(stat = "identity", 
           position = position_dodge(width = 0.9)) +
  labs(x = NULL, 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth", 
       fill = NULL) +
  scale_y_continuous(labels = scales::comma) +
  geom_text(aes(label = netWorth),
            position = position_dodge(width = 0.9),
            angle = 90,
            hjust = 1.1,
            color = "white") +
  theme_fivethirtyeight() +
  theme(axis.text.x = element_text(angle = 0, 
                                   hjust = 0.5, 
                                   vjust = 0.7))

```

Save the file to your hard disk

```{r eval=FALSE}
ggsave("Top_10_Billionaires.png",
       width = 9,
       height = 6,
       dpi=900)
```

```{r}
ggsave("Top_10_Billionaires.pdf",
       width = 9,
       height = 6,
       dpi=900)
```


Are you nostalgic about Excel graphs?

```{r}
ggplot(top10, aes(x = name, 
                  y = netWorth, 
                  fill = year2)) +
  geom_bar(stat = "identity", 
           position = position_dodge(width = 0.9)) +
  labs(x = "Name", 
       y = "Net Worth in Million $",
       title = "Top 10 Billionaires' Net Worth", 
       fill = NULL) +
  scale_y_continuous(labels = scales::comma) +
  geom_text(aes(label = netWorth),
            position = position_dodge(width = 0.9),
            angle = 90,
            hjust = 1.1,
            color = "white") +
  theme_excel() +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7))

```



## Original Graph


```{r}
ggplot(top10, aes(x = name2,
                  y = netWorth,
                  fill = year2)) +
  geom_bar(stat = "identity",
           position = position_dodge(width = 0.9)) +
  geom_text(aes(label = scales::comma(netWorth)),
            angle = 90,
            position = position_dodge(width = 0.9),
            hjust = 1.2,
            color = "white") +
  theme_fivethirtyeight() +
  scale_y_continuous(labels = scales::comma) +
  xlab("Name") +
  ylab("Net Worth in Million $") +
  guides(fill = guide_legend(title = NULL)) +
  ggtitle("Net Worth in Million $")

```



