---
title: "ggplot with image"
author: "Ashwin Malshé"
date: "06 October 2020"
output: html_document
---

For this exercise, we will install two new packages. Copy and paste the following command in your RStudio console and execute it.

install.packages(c('png', 'jpeg', "here", "showtext"))

Also, if you have never used `here` package before, install it as well:

install.packages("here")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(grid)
library(png)
library(jpeg)
library(here) # this package just points to the folder where you will save images
library(showtext)
```

## Use `annotation_custom` to include a `grob` or graphical object

Read the image from your computer. For this, you only need the file path.

In the code below I am reading a *jpg* file so I use the function `readJPEG` from package `jpeg`

```{r img1}
# Make sure that the file type is JPEG

img1 = readJPEG(here::here("Images", "diamond.jpg"))

# Next create a graphical object or a `grob`

grob1 = grid::rasterGrob(image = img1)
```

Create a plot using the above `grob`. In the plot `carat` is on x axis and `price` is on y. Map the color to `clarity`

For this, I am using `annotation_custom` function from `ggplot2`

```{r diamond}
ggplot(diamonds, aes(x = carat, y = price)) +
  annotation_custom(grob1,
                    xmin = -Inf, 
                    xmax = Inf, 
                    ymin = -Inf, 
                    ymax = Inf) +
  geom_point(aes(color = clarity),
             alpha = 0.4, # Adjust alpha to whatever you like
             show.legend = FALSE) +
  theme(panel.background = element_rect(fill = "white", # Choose the fill and color for panel
                                        color = "white"),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
  
```

## Using `annotation_raster` to avoid creating a `grob` separately

Import another image and create a bar graph.

```{r img2}
img2 = readJPEG(here::here("Images", "Bugatti.jpg"))
```

```{r bugatti}

ggplot(mpg, aes(class)) +
  annotation_raster(img2,
                    xmin = -Inf, 
                    xmax = Inf, 
                    ymin = -Inf, 
                    ymax = Inf,
                    interpolate = TRUE) +
  geom_bar(fill = "white",
           alpha = 0.1,
           color = "grey50") +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 70)) +
  labs(x = NULL, y = NULL)
```
```{r bugatti-2}

font_add("monoton", "Monoton-Regular.ttf")

showtext_auto()

ggplot(mpg, aes(class)) +
  annotation_raster(img2,
                    xmin = -Inf, 
                    xmax = Inf, 
                    ymin = -Inf, 
                    ymax = Inf,
                    interpolate = TRUE) +
  geom_bar(fill = "white",
           alpha = 0.1,
           color = "grey50") +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 70)) +
  labs(x = NULL, y = NULL,
       title = "Bugatti") +
  theme(plot.title = element_text(family = "monoton"))
```

```{r}
ggsave(here::here("Images", "bugatti_barplot.png"), width = 11, height = 8.5, dpi = 600)
```

