---
title: "Premade color palettes"
author: "Ashwin Malshe"
date: "10/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
theme_set(theme_minimal()) # this will set the default theme to theme_minimal
```

Although you can create your own color palettes in R or pass on your own colors to `ggplot2` using `scale_color_manual()` and `scale_fill_manual()`, there are also a few packages that already have nice color palettes ready to use out of the box.

Here I give examples of two such palettes.

## `wesanderson`

The first set of palettes comes from [Wes Anderson](https://en.wikipedia.org/wiki/Wes_Anderson) movies!

You can install the package from CRAN by typing this in the console:

install.packages("wesanderson")

The package documentation is available here:
https://github.com/karthik/wesanderson

Let's try it out.

**Warning**: Most of the palettes in this package have only 4-5 discrete colors. This means that if you have more than 4-5 levels in a variable, you can't use these palettes.

```{r}
library(wesanderson)
```

Import the palette you want and assign to an object.

```{r}
pal1 <- wes_palette("BottleRocket1")
```


```{r}
ggplot(mpg, aes(cty, hwy)) +
  geom_point(aes(color = class)) +
  scale_color_manual(values = pal1)
```

Exercise for you - 

Recreate the above plot using `wes_palette("Rushmore1")`. It has only 5 colors and `class` variable in `mpg` has sevel levels. How will you tackle that?


```{r}
pal2 <- wes_palette("Darjeeling1")
```

```{r}
ggplot(mpg, aes(cty, hwy)) +
  geom_point(aes(color = class)) +
  scale_color_manual(values = c(pal2, "#a3be8c", "#0063c4"))
```

## `viridis`

There are many great developers in the Rstats community. One of them is Bob Rudis. He co-created `viridis` package which has so far some of the best color palettes.

Install the package from CRAN by typing this in the console:

install.packages("viridis")

The package documentation is available here:
https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html


```{r}
library(viridis)
```
By default this package assumes a continuous color scale. That's why for discrete colors use `discrete = TRUE` argument.


```{r}
ggplot(mpg, aes(cty, hwy)) +
  geom_point(aes(color = class)) +
  scale_color_viridis(discrete = TRUE, direction = 1,
                      option = "viridis")
```


```{r}
ggplot(mpg, aes(cty, hwy)) +
  geom_point(aes(fill = class, color = class), 
             shape = 21, alpha = 0.5, size = 4) +
  scale_color_viridis(discrete = TRUE) +
  scale_fill_viridis(discrete = TRUE)
```
