---
title: "ggThemeAssist"
author: "Ashwin Malshe"
date: "10/06/2019"
output: html_document
---

Install `ggThemeAssist` from CRAN. Please DON'T put the next two lines in an R chunk. You need to do this only once. So, just copy and paste each line below in your RStudio console and execute it. Then restart RStudio.

install.packages("ggThemeAssist")
library(ggThemeAssist)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Create a simple plot and then enhance using `ggThemeAssist`

```{r}
p1 <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(method = "lm", se = FALSE)
```

Now highlight `p1` below and then select "ggplot Theme Assistant" from "Addins" button above.

```{r message=FALSE, warning=FALSE, error=FALSE}
p1 + theme(panel.grid.major = element_line(colour = "azure3"), 
    panel.grid.minor = element_line(colour = NA), 
    panel.background = element_rect(fill = NA)) +labs(x = "Displacement", y = "Highway Miles Per Gallon")
```



